---
title: "Sollers Background Module: Basic Statistics Session"
output: html_notebook
---

#R Studio Setup
1. Use the file drop down menu to create a new project
2. Go to Kaggle and download the Kiva crowd sourcing data set and put it in a *data*
sub directory.
3. Use File > New to open a new notebook
4. Use File > New to open a new R script.  The the script library. Put your library
commands in this file and source it in your notebooks.

#Data Structure in R.
The following shows some of the R data structure

This is an atomic vector of real (double) numbers.
```{r}
x <- c( 1, 3, 5, 5, 6, 7, 9, 10)
```

Sort a vector using standard R.
```{r}
y <- x[order(x)]
y

```
Index an element of a vector.
```{r}
y <- x[4]
```

```{r}
y
```
The properties of an object are length, typeof and attributes.
```{r}
length(y)
```
Index using a vector.
```{r}
y[c(4,5)]
```
```{r}
x[c(4,5)]
```
Make the methods of package ggplot2 available.
```{r}
library(ggplot2)
```

Create a data frame.
```{r}
d <- data.frame(x=x ) 
d
```
plot a density curve.  A density curve is like a histogram.
```{r}
ggplot(aes(x),data=d) +
  geom_density()
```
The sample variance is a function of the sum of the deviance from the mean.
```{r}
r <- (x - mean(x))^2
```
```{r}
sum(r)/(length(x) - 1)
```
Get help on the function rnorm.
```{r}
?rnorm
```

##Look at the use of dnorm, pnorm, qnorm and rnorm.

Simulate random normal variances.
```{r}
x <- rnorm(10000)
print(x[1:20])
d <- data.frame(x=x)
```
Create a normal curve using the normal random variables.
```{r}
ggplot(aes(x),data=d) +
  geom_density()
```
Explore the use of dnorm.

Create a grid of X values.
```{r}
s<-seq(-3,3,by=.001)
```


```{r}
d2 <- data.frame(x = s,y=dnorm(s))
d2[1:10,]
```
Create a box plot.  Usually there is a grouping variable.
```{r}
ggplot(d2) +
  aes(x=x,y=y) + 
  geom_boxplot()  
```
Slope in simple linear regression
slope = $r * \frac{\sigma_y}{\sigma_x} $
That means that the slope is related to the correlation between x and y.
The error term is many times assumed to be normally distributed.

Here is a norm distribution calculated using dnorm. 
d is for distribution. d stands for distribution.
```{r}
ggplot(d2) +
  aes(x=x,y=y) + 
  geom_point()  +
  geom_line()
```



What is pnorm used for? 
The area under the curve is 1. 
The area to the left of x=0 is 0.5
```{r}
pnorm(0)

```
The top of the curve is concave down.  The tails are concave up.
The two points where the curve changes from being concave up to 
concave down are called the inflection points.  
There are at -1 and 1 on the standard normal curve (mean = 0 and variance = 1).
```{r}
pnorm(c(-1,1))
```
The area between the two inflection points is .68.
```{r}
pnorm(1) - pnorm(-1)
```
What does qnorm calculate.
At what point on the x axis is the area to the left equal to 0.158?
```{r}
qnorm(0.158)
```

At what point is the area 0.841?
```{r}
qnorm(0.841)
```

At what x point is the area to the left .95 and the area to the right 0.05?
This is the critical value for a one tailed hypothesis test with a
significance level of 0.05.
```{r}
qnorm(.95)
```
What are the critical values for a two tailed hypothesis test with 
a significance level (area in both tails) of 0.05. 
```{r}
qnorm(c(0.025,0.975))
```
```{r}
qnorm(c(0.25,0.5,0.75))
```
Maximum likelihood.  
If trials are independent then their individual probabilities can 
be multiplied to get a join probability.
The probability of the sequence of HHHT.
Maximum likelihood select the value of p to make this function as 
large as possible.
Look at the values. The likelihood function is maximized when p=0.75.
This makes sense because 3/4 = 0.75.
```{r}
p <- .5
likelihood_f <-function(p){
  p*p*p*(1-p)
}
search_grid <- seq(0,1,by=0.01)

p_search<-data.frame(p = search_grid, likelihood_value = likelihood_f(search_grid))
p_search
                     
```


