---
title: 'Background Module: Basic Statistics'
output:
  pdf_document: default
  html_notebook: default
---

# Types of Data

There are two categories of measurements or data: 
*Categorical 
*Numerical or Qualitative and Quantitative.

Categorical data can be conceptualized as
tags.  In the English language, they are adjectives.  They describe a person, place or thing.  Examples are blue eyes, male or female people.  These characteristics are usually associated with counts of items.  They can be coded mathematically as 1 indicating the characteristic or 0 indicating absence of the characteristic.  Sometimes absence is coded with a -1 so the two codings add up to zero.  When there are only two levels, such as sex, then it is called binomial data, when there are many levels, such as color, it is considered nominal data, when the levels can be ordered, such as education level( BS, MS, PhD), it is called ordinal.

Numerical data on the other hand, is some type of measurement.  It could be the age of a person, distance, diastolic blood pressure, time after a certain event or temperature.  

Some measurements are counts of whole items.  These counts are discrete measurements.  Discrete means that for each measurement there is not another arbitrarily close measurement.  For example, if you are counting people, 100 is a possible number, but 100.5 is not.  All the counts are separated by one unit.  Other than count data, can you think of any other 
measurement that is discrete?

Other measurements can be measured with arbitrarily small precision.  This are interval numbers.  Weight can be measured 
to an arbitrarily small precision.  

Assuming the discrete data is ordered (if not it would be categorical) and there are many levels, it is generally 
analysed as continuous.  For example, money is discrete, but analyzed as continuous.  A likert scale (1 to 5) using in 
surveys is discrete, but sometime analyzed as continuous.

There are two other ways of looking at data: subjective vs. objective.  Many times qualitative data is subjective.  How do you feel as measured on a likert scale is subjective.  How long did it take you to walk four blocks is objective. 

# Representaion of data in R.
The computer language R uses vectors and lists to represent data.  These vectors and lists can be be give structure such as a matrix or a data table. The data table is similar to an Excel table.  A matrix is what you learned about in finite math.

The elements of the vector or list are called types.  The four main types are be logical, integer, double or character.  

```{r}
x_logical <- c(TRUE,FALSE,TRUE)
typeof(x_logical)
```
The function class can also be used.
```{r}
class(x_logical)
```
Some data structures consist of a vector of elements that are of the same type.  These are called atomic vectors.

```{r}
x_atomic <- c(1,2,45,69)
is.atomic(x_atomic)
```
```{r}
x_list <- list(1,2,45,69)
is.atomic(x_list)
```
In R, there is something called and atomic vector, it is created with the operator c.
```{r}
y_atomic <- c(1,2,3,"A","B")
is.atomic(y_atomic)
```
To get the vector to be atomic, R does some automatic type conversion.  Here everything is converted to character.
```{r}
y_atomic
```
Lists don't have to be atomic.
```{r}
y_list <- list(1,2,3,"A","B")
print(y_list)
```
Structure to data sets is created using tributes.  
```{r}
x_matrix <- matrix(1:12,nrow = 4)
x_matrix
```
We can see the attributes
```{r}
attributes(x_matrix)
```
The class is set to matrix.
```{r}
class(x_matrix)
```
A data frame is a list of data columns


```{r}
list(letters[1:10])
x_data_frame <- data.frame(x = 1:10, y=letters[1:10],z = list(letters[1:10]),stringsAsFactors = FALSE)
x_data_frame
```
```{r}
x_data_frame[[1]]
```

```{r}
typeof(x_data_frame[[1]])
```
```{r}
x_data_frame[[2]]
```

```{r}
typeof(x_data_frame[[2]])
```
```{r}
x_data_frame[[3]]
```
```{r}
typeof(x_data_frame[[3]])
```
```{r}
is.atomic(x_data_frame[[3]])
```
This data table consists of a list of atomic vectors.

```{r}
typeof(x_data_frame)
```

# Statistical Inference

Gain information on a population from a sample.  The sample must represent the population.  Gather a 
sample and learn about the population from the sample.  

Things to learn
*Differentiate between a  population and a sample?
*Why is it important to have a random sample?
*What is inference?
*What is a parameter?
*What is a statistic?  How can it be used to estimate a parameter.

# Two Branches of Statistics
What is the difference between a Frequentest and a Bayesian?
The data is random or the parameter is random.
Random or uncertainty.

# Descriptive Statistics
Things to know 
*Measures of Central Tendency
*Measures of dispersion

How are they used to estimate population parameters?

Understand how a distribution of numbers is represented by a histogram or density plot.

What is skew, how does it affect the measures of central tendency?

What are good plots to show dispersion?
*box plot
*histogram

What are measures that show dispersion
*variance
*percentiles
*Quartiles
*Basic dispersion of the normal distribution

## Association between two variables.
*Covariance
*Correlation

#Proability Distributions
*Continuous 
*Discrete

What is the cumulative distribution.

What are:
*probability
*quantile
*density
*random variable

# Terms used to desribe data
*Stochastic process
*Identically distributed
*Independent
*Independent and identically distributed
*dependent
*Mutually exclusive

# Seven Concepts
1. Errors and Residuals
1. Point Estimates
1. Standard Error of the mean
1. Sampling Error
1. Confidence Interval
1. Confidence Levels
1. Hypothesis Testing.
















